---
title: "R Notebook"
output: html_notebook
---

**Explanation:**
- This R Notebook chunk is labeled "Download and Read Dataset" for clarity.
- The `echo=TRUE` argument is used to include the code in the output.
- The code checks if the 'downloader' package is installed and installs it if not.
- The 'downloader' package is loaded.
- The dataset is downloaded from the specified URL and saved with the filename "mice_pheno.csv".
- The dataset is then read into a data frame named 'dat' using the `read.csv` function.


```{r Download and Read Dataset, echo=TRUE}
# Load the 'downloader' package if not already installed
if (!requireNamespace("downloader", quietly = TRUE)) {
  install.packages("downloader")
}
library(downloader)

# Download the dataset
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/mice_pheno.csv"
filename <- "mice_pheno.csv"
download(url, destfile = filename)

# Read the dataset into a data frame
dat <- read.csv(filename)

dat%>%skim()

```


**Explanation:**
- This R Notebook chunk is labeled "Data Filtering and Descriptive Statistics" for clarity.
- The `echo=TRUE` argument is used to include the code in the output.
- The required 'dplyr' library is loaded.
- The dataset is filtered to create two populations based on specified criteria.
- The means of the filtered populations are calculated.
- The difference in means is printed.

```{r Data Filtering and Descriptive Statistics, echo=TRUE}
# Load the required libraries
library(dplyr)

# Filter the dataset to create populations based on criteria
controlPopulation <- filter(dat,Sex == "F" & Diet == "chow") %>%  
  select(Bodyweight) %>% unlist
hfPopulation <- filter(dat,Sex == "F" & Diet == "hf") %>%  
  select(Bodyweight) %>% unlist
# Calculate means of the populations
mu_hf <- mean(hfPopulation)
mu_control <- mean(controlPopulation)

# Print the difference in means
print(mu_hf - mu_control)
```


**Explanation:**
- This R Notebook chunk is labeled "Variance Calculation and Comparison" for clarity.
- The `echo=TRUE` argument is used to include the code in the output.
- The population variance is calculated using the formula \( \frac{1}{N} \sum_{i=1}^{N} (x_i - \bar{x})^2 \), where \( N \) is the length of the population.
- The code checks if the variance calculated using the `var()` function is identical to the calculated population variance.
- The code also checks if the adjusted variance (\( \frac{N-1}{N} \times \text{var}(x) \)) is identical to the calculated population variance.

```{r Variance Calculation and Comparison, echo=TRUE}
# Assign controlPopulation to variable x
x <- controlPopulation

# Get the length of x
N <- length(x)

# Calculate population variance
populationvar <- mean((x - mean(x))^2)

# Check if the variance calculated using var() is identical to the population variance
identical(var(x), populationvar)

# Check if the adjusted variance is identical to the population variance
identical(var(x) * (N - 1) / N, populationvar)

library(rafalib)
sd_hf <- popsd(hfPopulation)
sd_control <- popsd(controlPopulation)

```

**Explanation:**
- This R Notebook chunk is labeled "Sampling and Simulation" for clarity.
- The `echo=TRUE` argument is used to include the code in the output.
- A sample size \(N\) is set for the hf and control populations.
- Random samples of size \(N\) are created from hfPopulation and controlPopulation.
- A vector `Ns` is defined, representing different sample sizes for simulations.
- The number of simulations (\(B\)) is set to 10000.
- Simulations are performed using the `sapply` function, calculating the differences in means for each sample size and replicating the process for 10000 times. The results are stored in the `res` matrix.

```{r Sampling and Simulation, echo=TRUE}
# Set the sample size for hf and control populations
N <- 12

# Create random samples from hfPopulation and controlPopulation
hf <- sample(hfPopulation, N)
control <- sample(controlPopulation, N)

# Define a vector of different sample sizes (Ns)
Ns <- c(3, 12, 25, 50)

# Set the number of simulations (B)
B <- 10000

# Perform simulations to calculate differences in means for each sample size
res <- sapply(Ns, function(n) {
  replicate(B, mean(sample(hfPopulation, n)) - mean(sample(controlPopulation, n)))
})


```

**Explanation:**
- This R Notebook chunk is labeled "QQ Plots for Simulated Differences in Means" for clarity.
- The `echo=TRUE` argument is used to include the code in the output.
- A custom function `mypar` is defined to set the layout using `par` for a 2x2 plot.
- The layout is set using `mypar(2, 2)`.
- A loop is used to iterate through different sample sizes (`Ns`).
- For each sample size, the average and standard deviation of the simulated differences in means are calculated for titles.
- QQ plots are created using `qqnorm` and `qqline` to visualize the distribution of the simulated differences in means. Each plot corresponds to a different sample size.


```{r QQ Plots for Simulated Differences in Means, echo=TRUE}
# Define a custom par function (mypar)
mypar <- function(nr, nc) {
  par(mfrow = c(nr, nc), mar = c(4, 4, 2, 1))
}

# Set the layout for a 2x2 plot
mypar(2, 2)

# Loop through different sample sizes
for (i in seq(along = Ns)) {
  # Calculate average and standard deviation for titles
  titleavg <- signif(mean(res[, i]), 3)
  titlesd <- signif(sd(res[, i]), 3)
  title <- paste0("N=", Ns[i], " Avg=", titleavg, " SD=", titlesd)

  # Create QQ plots
  qqnorm(res[, i], main = title)
  qqline(res[, i], col = 2)
}
```

**Explanation:**
- This R Notebook chunk is labeled "QQ Plots for Simulated t-Statistic" for clarity.
- The `echo=TRUE` argument is used to include the code in the output.
- A custom function `mypar` is defined to set the layout using `par` for a 2x2 plot.
- The sample sizes (`Ns`) and the number of simulations (`B`) are set.
- A function `computetstat` is defined to compute a t-statistic for given sample sizes.
- Simulations are performed using `sapply` to calculate t-statistics for each sample size and replicate the process for 10000 times. The results are stored in the `res` matrix.
- QQ plots are created using `qqnorm` and `qqline` to visualize the distribution of the simulated t-statistics. Each plot corresponds to a different sample size.
```{r QQ Plots for Simulated t-Statistic, echo=TRUE}
# Define a custom par function (mypar)
mypar <- function(nr, nc) {
  par(mfrow = c(nr, nc), mar = c(4, 4, 2, 1))
}

# Set the sample sizes (Ns) and number of simulations (B)
Ns <- c(3, 12, 25, 50)
B <- 10000

# Function to compute a t-statistic
computetstat <- function(n) {
  y <- sample(hfPopulation, n)
  x <- sample(controlPopulation, n)
  (mean(y) - mean(x)) / sqrt(var(y) / n + var(x) / n)
}

# Perform simulations to compute t-statistics for each sample size
res <- sapply(Ns, function(n) {
  replicate(B, computetstat(n))
})

# Set the layout for a 2x2 plot
mypar(2, 2)

# Loop through different sample sizes
for (i in seq(along = Ns)) {
  # Create QQ plots for t-statistics
  qqnorm(res[, i], main = Ns[i])
  qqline(res[, i], col = 2)
}

```



































