---
title: "R Notebook_Population, Samples, and Estimates Exercises"
output: html_notebook
---
Import data into R

```{r}
library(downloader) 
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/mice_pheno.csv"
filename <- basename(url)
download(url, destfile=filename)
dat <- read.csv(filename)
dat<-na.omit(dat)
head(dat)

```
1. Use dplyr to create a vector x with the body weight of all males on the control (chow) diet. What is this population’s average?


```{r}
library(dplyr)
x<-filter(dat,Sex=="M" & Diet=="chow")$Bodyweight
mean_x<-mean(x)
print(mean_x)
```
2. Now use the rafalib package and use the popsd function to compute the population standard deviation.

```{r}
library(rafalib)
sd_x<-popsd(x)
sd_x
```
3. Set the seed at 1. Take a random sample X of size 25 from x. What is the sample average?
```{r}
set.seed(1)
control_x<-sample(x,25)
control_x_mean<-mean(control_x)
control_x_mean
```

4. Use dplyr to create a vector y with the body weight of all males on the high fat (hf) diet. What is this population’s average?
```{r}
library(dplyr)
y<-filter(dat,Sex=="M" & Diet=="hf")$Bodyweight
mean_y<-mean(y)
print(mean_y)
```
5. Now use the rafalib package and use the popsd function to compute the population standard deviation.

```{r}
library(rafalib)
sd_y<-popsd(y)
sd_y
```
5. Set the seed at 1. Take a random sample Y of size 25 from y. What is the sample average?
```{r}
set.seed(1)
control_y<-sample(y,25)
control_y_mean<-mean(control_y)
control_y_mean
```
7. What is the difference in absolute value between y¯−x¯ ?
 

```{r}
abs_diff_x_y_bar <- abs(mean_y - mean_x - (control_x_mean - control_y_mean))
print(abs_diff_x_y_bar)
```
Repeat the above for females. Make sure to set the seed to 1 before each sample call. What is the difference in absolute value between y¯−x¯ ?
```{r}
set.seed(1)

x_female <- filter(dat, Sex == "F" & Diet == "chow")$Bodyweight
population_mean_x_female <- mean(x_female)
population_sd_x_female <- popsd(x_female)

sample_X_female <- sample(x_female, 25)
sample_mean_X_female <- mean(sample_X_female)

y_female <- filter(dat, Sex == "F" & Diet == "hf")$Bodyweight
population_mean_y_female <- mean(y_female)
population_sd_y_female <- popsd(y_female)

sample_Y_female <- sample(y_female, 25)
sample_mean_Y_female <- mean(sample_Y_female)


abs_diff_x_y_bar_female <- abs(population_mean_y_female - population_mean_x_female - (sample_mean_X_female - sample_mean_Y_female))
print(abs_diff_x_y_bar_female)

```

 























