---
title: "R Notebook_Random Variables"
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---

Importing Data into R

```{r}
library(devtools)
install_github("genomicsclass/dagdata",force=TRUE)
dir<-system.file(package="dagdata")
list.files(dir)
list.files(file.path(dir,"extdata"))
filename<-file.path(dir,"extdata/femaleMiceWeights.csv")
dat <- read.csv(filename)
View(dat)
head(dat)
```

```{r}
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/femaleControlsPopulation.csv"
filename_1<- "femaleControlsPopulation.csv"
if (!file.exists(filename_1)) download(url,destfile=filename_1)
population<-read.csv(filename_1)
population<-unlist(population)
head(population)
```

Calculate Mean Bodyweights for Control and Treatment Groups

```{r}
install.packages("dplyr")
library(dplyr)
control<-filter(dat,Diet=="chow")%>%select(Bodyweight)%>%unlist
treatment<-filter(dat,Diet=="hf")%>%select(Bodyweight)%>%unlist
print(mean(control))
print(mean(treatment))
obsdiff<-mean(treatment)-mean(control)
print(obsdiff)
```

Random variables

```{r}
View(population)
head(population)
control_population<-sample(population,12)
mean(control_population)
control_population_1<-sample(population,12)
mean(control_population_1)

```

The NULL Hypothesis
Now, let's revisit our calculated average difference, obsdiff. As scientists, it's important for us to maintain a skeptical mindset. We should question whether this observed difference can be attributed solely to the variations in diet. To address this, we might wonder what would happen if we were to provide the same diet to all 24 mice. Would we still expect to observe a difference as significant as the one we've calculated? Statisticians refer to this scenario as the null hypothesis. The term "null" serves as a reminder that we are approaching the question with skepticism, acknowledging the possibility that there may be no discernible difference.

Since we have access to the entire population, we can actually observe as many instances as we wish where the diet has no impact. We can achieve this by randomly selecting 24 control mice, subjecting them to the same diet, and then recording the difference in means between two randomly divided groups of 12 each.
```{r}
control_1<-sample(population,12)
treatment_1<-sample(population,12)
print(mean(treatment_1)-mean(control_1))


n<-10000
null<-numeric(n) 

for (i in 1:n) {
  control_2<-sample(population,12)
  treatment_2<-sample(population,12)
  null[i]<-mean(treatment_2)-mean(control_2)
}
percentage_bigger_than_obsdiff<-mean(null>=obsdiff)*100
```
A negative value for percentage_bigger_than_obsdiff indicates that a relatively small proportion of the values in the "null" distribution are greater than or equal to obsdiff. This means that our observed difference, obsdiff, is on the lower end of the distribution of differences when the diet has no effect. In other words, it's not a common occurrence in the simulated scenarios where there's no difference in the diet's effect on the groups.

This result is consistent with hypothesis testing in statistics. A small percentage (close to 0) indicates that the observed difference is not likely to have occurred by random chance when the null hypothesis is true (i.e., when there's no effect of the diet). It suggests that the observed difference is statistically significant and likely a result of the diet.

Distributions
```{r}
install.packages("UsingR")
library(UsingR)
x<-father.son$fheight
round(sample(x,10),1)
```
Cumulative Distribution Function

CDF: Cumulative Distribution Function
ECDF: Empirical CDF
```{r}
smallest<-floor(min(x))
largest<-ceiling(max(x))
values<-seq(smallest,largest,len=300)
height_ecdf<-ecdf(x)
plot(values, height_ecdf(values), type="l",
     xlab="a (Height in inches)",ylab="Pr(x <= a)")
```
Histograms
```{r}
hist(x)
bins<-seq(smallest,largest)
hist(x,breaks=bins, xlab="Height (in inches)",main="Adult men heights")
```

Probability Distribution
```{r}
n<-100
library(rafalib)
nullplot(-5,5,1,30, xlab="Observed differences (grams)", ylab="Frequency")
totals<-numeric(11)
for (i in 1:n) {
  control_3<-sample(population,12)
  treatment_3<-sample(population,12)
  nulldiff<-mean(treatment_3)-mean(control_3)
  j<-pmax(pmin(round(nulldiff)+6,11),1)
  totals[j]<-totals[j]+1
  text(j-6,totals[j],pch=15,round(nulldiff,1))
}
```
```{r}
hist(null,freq=TRUE)
abline(v=obsdiff,col="yellow",lwd=2)
```




















